\sysfull

### 2017-08-10 paper/relatedwork.tex
finished the pass over related work and added the VM stuff

\noindent \textbf{Intermittently-Powered Devices:} There is plentiful of research results available on energy harvesting, or more specifically wireless power provision, for embedded devices like sensors or microcontrollers. This paragraph is too short to cover this area extensively and we refer the reader to overview papers~\cite{prasad_comst_2014,sample_procieee_2013,huang:commag:2015,visser_procieee_2013,kamalinejad_commag_2015,ku_cst_2016} for in-depth discussion. With power consumption of electronics going down dramatically with every year, computing systems can be now powered completely from ambient energy. This opens new research areas such as radio frequency-powered computers~\cite{patel_pervasive_2017,rf_powered_computing_gollakota_2014}. Indeed, plenty of new embedded platforms appear that are powered solely from ambient (intentional or non-intentional) energy and include~\cite{wisp5,moo,zhao_rfid_2015,holleman_biocas_2008,thomas_jbcs_2012,naderiparizi_rfid_2015,rodriguez_tbcs_2015,liu_sigcomm_2013,kicksat,nadeau_naturebio_2017}. But there is a price tag associated to them: harvested energy is miniscule, unstable, difficult to predict ergo \emph{intermittent}. Therefore, to guarantee completion of computing, programming and runtime models are being actively researched that are specifically tailored to these intermittently-powered systems. \todo{Revise this section}{Brandon}

\noindent \textbf{Intermittent Execution Environments:} Checkpoint runtimes include DINO~\cite{dino}, Hibernus/Hibernus++~\cite{hibernus,hibernusplusplus}. The only available Task-based intermittent runtime systems are Chain~\cite{chain} and Alpaca~\cite{alpaca}. \sys goes beyond what any task-based system performed by providing \emph{automatic} task creation and \emph{adaptivity} to energy state and energy storage \emph{obliviousness} of the compiled code. \todo{complete this section}{Przemek}

\noindent \textbf{Memory Virtualization:} Review of memory virtualization environments (not necessarily for intermittent power devices). \todo{Fill-in this section}{Przemek}

>>>>

\noindent \textbf{Intermittently-Powered Devices:} There is a large body of
research on energy harvesting and wirelessly powered embedded devices like
sensors or microcontrollers. Several recent reviews provide a broad overview
papers~\cite{prasad_comst_2014,sample_procieee_2013,huang:commag:2015,visser_procieee_2013,kamalinejad_commag_2015,ku_cst_2016}
in depth.  As computing power consumption continues to decrease, a new wave of
{\em ambiently} powered systems is emerging, powered by radio
waves~\cite{patel_pervasive_2017,rf_powered_computing_gollakota_2014}.
Radio-powered
devices~\cite{wisp5,moo,zhao_rfid_2015,holleman_biocas_2008,thomas_jbcs_2012,naderiparizi_rfid_2015,rodriguez_tbcs_2015,liu_sigcomm_2013,kicksat,nadeau_naturebio_2017}
are realted to \sys because they are the technological foundation of an
important category of intermittent computing devices. The emergence of
intermittent computing hardware platforms has led to the development of tooling
and instrumentation that support using these platforms for system
building~\cite{hester_sensys_2014,hester_sensys_2015,edb,stork,wisent}.

\noindent \textbf{Intermittent Execution Environments:} 
The proliferation of energy-harvesting hardware led to a need for system
support.  Early work developed runtime environments for energy-harvesting
systems assuming simple, small computations that do not exceed a predictable
burst of energy~\cite{dewdrop} or long power-cycles and priority-based
scheduling to ensure important tasks execute often
enough~\cite{sorber_sensys_2007}.  These early efforts did not consider
intermittence or data consistency issues. 
%
Later efforts supported intermittent computing using dynamic, on-demand
checkpoints include Mementos~\cite{mementos}, Quickrecall~\cite{quickrecall},
Hibernus/Hibernus++~\cite{hibernus,hibernusplusplus} and
HarvOS~\cite{mottola2017harvos}. Dynamic checkpointing systems may checkpoint
at any point, making it difficult to implement application-level atomicity
requirements, such as collecting correlated sensor samples together.  Moreover,
some of these systems require hardware support for a device to measure its
stored energy, which requires additional hardware and may incur an energy cost. 
Moreover, these systems do not checkpoint non-volatile state, compromising 
memory consistency.
%
DINO~\cite{dino} selectively versions non-volatile memory and checkpoints
volatile state, ensuring progress and memory consistency.
Rachet~\cite{ratchet} assumes all memory is non-volatile and checkpoints
volatile state.  Clang~\cite{hicks_isca_2017} takes a Ratchet-like approach
that leverages dedicated hardware to eliminate checkpointing and versioning
overhead.

Two task-based system developed in recent work are Chain~\cite{chain} and
Alpaca~\cite{alpaca}.  Using static tasks, these systems eliminate the need to
checkpoint volatile state.  Using novel, channel-based memory
models~\cite{chain} or automatic privatization and redo-logging~\cite{alpaca}
these systems avoid checkpointing overheads, do not require fully non-volatile
memory, and require no special hardware.  Moreover, task-based models
facilitate specifying application-level atomicity properties.

\sys relates to earlier efforts on intermittent computing in their shared
purpose of making it simple to write correct programs that run intermittently,
but is distinct in its approach and mechanism.  \sys relates more closely to
task-based intermittent programming execution models because \sys also relies
on statically defined tasks to avoid the need to checkpoint volatile state.
\sys's automatic compilation spares the programmer the work of writing
task-based code that prior systems demand.  \sys's memory virtualization
provides an alternative design for ensuring memory consistency that is
optimized for bulk accesses to task-shared data with high locality.  \sys's
task coalescing optimization ameliorates the key overheads associated with
committing tasks present in these prior systems.

{\noindent \bf Memory Virtualization}

Prior work on operating system support for embedded systems and sensor motes
has studied a variety of memory virtualization strategies that relate to \sys's
memory virtualization mechanism.

Several operating system efforts developed support for memory management.
TinyOS and nesC support dynamic memory management~\cite{nesc}.  Later work
extended the memory manager to support memory virtualization backed by flash
memory~\cite{sensornetvm} and to ensure memory and type
safety~\cite{tinyosmemorysafety}. SOS~\cite{sos}, Contiki~\cite{contiki}, and
t-kernel~\cite{tkernel} also developed memory management abstractions that
virtualize memory size and provide safe and indirect access.
Mat\'e~\cite{mate} developed full virtual machines support for sensor nodes,
virtualizing not just memory resources, but other state and peripheral
resources.  These prior efforts to virtualize memory in embedded and sensor
node systems are similar in their purpose and mechanism to \sys, but \sys is
distinct in its goal.  The goal of \sys is to provide consistent, intermittent
execution, leveraging the performance benefits of efficient bulk copying.  In
contrast, these prior efforts were focused more on programmability and run time
reliability properties provided by virtual memory.

Another prior effort that is related to \sys is work on unbounded, page-based
transactional memory and deterministic parallel runtime
systems~\cite{pagebasedtm,grace}.  These prior efforts have a different
mechanism and purpose than \sys -- ensuring that data are consistent and
deterministically updated during concurrent executions.  What makes these
efforts related is that they manage state required to ensure consistency at the
granularity of pages to amortize checking and tracking costs.  \sys also tracks
accesses at the granularity of a page and uses its paging mechanism to ensure
consistency.  Moreover, \sys's paging implementation, which keeps a shadow page
for each page to use during commit is very similar to the shadow paging scheme
use for transactional commit~\cite{pagebasedtm}.
